---
title: "Cleaning Respiration Data Workflow"
author: "Nathan Brouwer"
date: "May 3, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
#plotting function
library(ggplot2)

#cleans up ggplot
library(cowplot)

#reshape data to calcualte minimum
```

## Source function

Load the 2 functions used

* clean_resp_dat()
* code_flushes()


```{r}
source('./R/clean_resp_dat.R')
source('./R/code_flushes.R')
```



## Worflow

### Steps in the workflow

* Use clean_resp_dat() function to load, clean, and re-save data
* Use code_flushes() to code when the "flush periods" are

Note that clean_resp_dat() is a full-service function; you give it a file folder and a file name, and it loads the data, cleans it, provides some summary output, and resaves it as csv.

### Arguments in clean_resp_dat() function

* raw_data_folder = where the raw data lives; current needs to have "/" in proper place.
* raw_data_file = the name of the exact file
* clean_datat_folder = where a .csv version of the data should be saved
* clean_file_name = what to call the clened file
* object_name = what to call the R object containing the clean memory; this is what will be subsequently used for analysis

### Load & clean data

First, load and clean the data.  This produes and object my_phys that will subsequently be used.

```{r}
clean_resp_dat(raw_data_folder = "./data-raw/",
                           raw_data_file = "23-apr.csv",
                           clean_data_folder = "./data/",
                           clean_file_name = "my_clean_data.csv",
                           object_name = "my_phys")
```


Confirm the presence and size of the my_phys object

```{r}
dim(my_phys)
```


### Code the flush periods

The code_flushes() function parses the cleaned data and designates which measurements were during "flush" periods and shoudl not be used in analysis, and which are valid data points ("mous").

* dat = R dataframe object
* time_col = column contain the time the data was collected
* variable_col = column with focal measurement variable
* channel_col = column of data frame with the channel (chamber) designation
* n_chambers = the number of channesl
* n_cycles = the numbr of full cycles through all channels; do not include partial cycles
* measures_per_cycle = number of times measurements were taken per cycle (? shoudl this be measures_per_channel)
* flush_cutoff = number of initial measurements within a cycle to consider part of the flush phase

```{r}
my_phys_coded <- code_flushes(dat = my_phys,
                         time_col = "time.min",
                         variable_col = "Q.S102.O2.Pcor",
                         channel_col = "channel",
                         n_chambers = 8,
                         n_cycles = 4,
                         measures_per_cycle = 48,
                         flush_cutoff = 19)
```


## Plot data

```{r}
ggplot(data = my_phys_coded,
       aes(y = Q.S102.O2.Pcor ,
           x = chamb.measure.i,
           color = factor(phase))) + 
  geom_point() +
  facet_wrap(~channel)
```


## Calcualte minimum

### Global minimum

Lowest value for a each chamber accross all cycles
```{r}
global.min <- dcast(data = my_phys_coded,
      formula = channel ~ .,
      value.var = "Q.S102.O2.Pcor",
      fun.aggregate = min)

names(global.min)[2] <- "min"
```


### Minimum by Phase

```{r}
phase.min <- dcast(data = my_phys_coded,
      formula = channel ~ phase,
      value.var = "Q.S102.O2.Pcor",
      fun.aggregate = min)

```


### Variance by Phase

```{r}
phase.var <- dcast(data = my_phys_coded,
      formula = channel ~ phase,
      value.var = "Q.S102.O2.Pcor",
      fun.aggregate = var)

```
